New insights into the Weddell Sea ecosystem applying a network approach

### 2022-11-09 Manuscript/Manuscript.Rmd
Starting w/ Discussion

Regarding the species' mean interaction strength and its relationship with the unweighted properties (Figure 4A-D), we found that interaction strength-trophic level showed a significant positive linear regression: the higher the trophic level of the species, the higher its mean interaction strength. We also found a significant but less evident positive relationship with degree (total number of interactions). On the other hand, there was no significant relationship with trophic similarity. Considering species habitat, the "Benthopelagic" and "Pelagic" categories presented the two species with the highest mean interaction strength, though "Demersal" and "Land-based" habitats concentrated the majority of the species with relatively higher interaction strength. Species inhabiting the benthic realm showed the lowest mean interaction strength (Figure 4D).

@@

It's important to note that since the proportion of Jacobians that are locally stable or QSS was zero for the Weddell Sea food web, we considered the mean maximum eigenvalue as the stability index. Our results showed that the majority of the species had no significant impact on the stability of the food web when extinct (Figure 5). This is shown in Figure 5, where most of the points lie around the zero value of the 'Stability difference' y axis, meaning that the stability for the food web minus a given species is similar to the stability for the whole network. However, there were few species that when become extinct altered significantly the stability (QSS) of the food web. Most of such species generated positive significant differences in the stability of the food web, which mean that they increased the network's stability when deleted. There were two exceptions, Pagetopsis macropterus (demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod), that when deleted decreased the stability. Overall, 15 out of 490 species (3.06%) gave rise to significant changes in the food web's stability (Table 2).

After exploring the stability difference against the species properties (Figure 5), we found that those species that generated a significant impact on the stability of the food web are characterized by: 1) high mean interaction strength; 2) mid to high trophic levels (TL > 3.2); 3) relatively high number of interactions (Degree > 25); and 4) mid to low trophic similarity values (TS < 0.16). Habitat wise, species with a significant impact on the stability are in all habitats, except for the benthic realm. Table 2 shows the results for such species.

@@

Dominance of weak interactions in concordance with other studies.

@@

"Low functional redundancy at key trophic levels makes these ecosystems (polar pelagic) particularly sensitive to change". [@Murphy2016]

@@

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Frequency distribution of interaction strengths for the Weddell Sea food web (n = 490)."}
knitr::include_graphics("Fig3_IntDist.png")
```

>>>>

Following the above results for the distribution of the interaction strengths, we found that the species mean interaction strength was different. We also found that such weighted property is related more or less with unweighted properties (Figure 4A-D). In this regard, interaction strength - trophic level showed a significant positive linear regression: the higher the trophic level of the species, the higher its mean interaction strength. We also found a significant but less evident positive relationship with degree or the total number of interactions. On the other hand, there was no significant relationship with trophic similarity. Considering species habitat, the "Benthopelagic" and "Pelagic" categories presented the two species with the highest mean interaction strength (Orcinus orca and Mesonychoteuthis hamiltoni, respectively), though "Demersal" and "Land-based" habitats concentrated the majority of the species with relatively higher interaction strength. Species inhabiting the benthic realm showed the lowest mean interaction strength (Figure 4D).

@@

It's important to note that since the proportion of Jacobians that are locally stable or QSS was zero for the Weddell Sea food web, we considered the mean maximum eigenvalue as the stability index. When performing the extinction simulations, our results showed that the majority of the species had no significant impact on the stability of the food web when extinct (Figure 5). This is shown in Figure 5, where most of the points lie around the zero value of the 'Stability difference' y axis, meaning that the stability for the food web minus a given species is similar (not significant) to the stability for the whole network. However, there were few species that when become extinct altered significantly the stability of the food web. Most of such species generated positive significant differences in the stability of the food web, which mean that they increased the network's stability when deleted. There were two exceptions, Pagetopsis macropterus (demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod), that when deleted decreased the stability. Overall, 15 out of 490 species (3.06%) gave rise to significant changes in the food web's stability (Table 2).

After exploring the stability difference against the species properties (Figure 5), we found that those species that generated a significant impact on the stability of the food web were characterized by: 1) high mean interaction strength; 2) mid to high trophic levels (TL > 3.2); 3) relatively high number of interactions (Degree > 25); and 4) mid to low trophic similarity (TS < 0.16). Habitat wise, species with a significant impact on the stability were present in all habitats, except for the benthic realm. Table 2 shows these results for such species.

@@

Principal results:
1) First study to
      * consider the interaction dimensionality to estimate interaction strength in a food web framework

2) Asymmetric distribution of IS at the food web level; dominance of weak interactions

3) Strong relationship between interaction strength and trophic level. Less stronger but evident relationship with degree

4) Few species are important in terms of stability: discuss species characteristics

1)
2)
Dominance of weak interactions in concordance with other studies.

@@

3)

4)
Our results showed that only few species are key in terms of stability in the Weddell Sea food web. This is in concordance with other studies of complex empirical food webs in high-latitude marine ecosystems [@Kortsch2015; @Marina2018; @Rodriguez2022]. Such key species are characterized by: 1) high mean interaction strength; 2) mid to high trophic level; 3) relatively high number of interactions; and 4) mid to low trophic similarity.

Our results contradict studies suggesting that mid-trophic level species are key in marine polar ecosystems.

The trophic similarity index might be interpreted as a proxy for the functional redundancy of the food web. In this sense, ...
Low functional redundancy at key trophic levels makes these ecosystems (polar pelagic) particularly sensitive to change. [@Murphy2016]

@@

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Frequency distribution of interaction strengths for the Weddell Sea food web. Total number of interactions = 16041."}
knitr::include_graphics("Fig3_IntDist.png")
```

### 2022-11-09 Manuscript/Manuscript.tex
Starting w/ Discussion

Regarding the species' mean interaction strength and its relationship
with the unweighted properties (Figure 4A-D), we found that interaction
strength-trophic level showed a significant positive linear regression:
the higher the trophic level of the species, the higher its mean
interaction strength. We also found a significant but less evident
positive relationship with degree (total number of interactions). On the
other hand, there was no significant relationship with trophic
similarity. Considering species habitat, the ``Benthopelagic'' and
``Pelagic'' categories presented the two species with the highest mean
interaction strength, though ``Demersal'' and ``Land-based'' habitats
concentrated the majority of the species with relatively higher
interaction strength. Species inhabiting the benthic realm showed the
lowest mean interaction strength (Figure 4D).

@@

It's important to note that since the proportion of Jacobians that are
locally stable or QSS was zero for the Weddell Sea food web, we
considered the mean maximum eigenvalue as the stability index. Our
results showed that the majority of the species had no significant
impact on the stability of the food web when extinct (Figure 5). This is
shown in Figure 5, where most of the points lie around the zero value of
the `Stability difference' y axis, meaning that the stability for the
food web minus a given species is similar to the stability for the whole
network. However, there were few species that when become extinct
altered significantly the stability (QSS) of the food web. Most of such
species generated positive significant differences in the stability of
the food web, which mean that they increased the network's stability
when deleted. There were two exceptions, Pagetopsis macropterus
(demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod),
that when deleted decreased the stability. Overall, 15 out of 490
species (3.06\%) gave rise to significant changes in the food web's
stability (Table 2).

After exploring the stability difference against the species properties
(Figure 5), we found that those species that generated a significant
impact on the stability of the food web are characterized by: 1) high
mean interaction strength; 2) mid to high trophic levels (TL
\textgreater{} 3.2); 3) relatively high number of interactions (Degree
\textgreater{} 25); and 4) mid to low trophic similarity values (TS
\textless{} 0.16). Habitat wise, species with a significant impact on
the stability are in all habitats, except for the benthic realm. Table 2
shows the results for such species.

@@

Dominance of weak interactions in concordance with other studies.

@@

``Low functional redundancy at key trophic levels makes these ecosystems
(polar pelagic) particularly sensitive to change''. \citep{Murphy2016}

@@

\begin{figure}
\includegraphics[width=12cm]{Fig3_IntDist} \caption{Frequency distribution of interaction strengths for the Weddell Sea food web (n = 490).}\label{fig:unnamed-chunk-3}
\end{figure}

>>>>

Following the above results for the distribution of the interaction
strengths, we found that the species mean interaction strength was
different. We also found that such weighted property is related more or
less with unweighted properties (Figure 4A-D). In this regard,
interaction strength - trophic level showed a significant positive
linear regression: the higher the trophic level of the species, the
higher its mean interaction strength. We also found a significant but
less evident positive relationship with degree or the total number of
interactions. On the other hand, there was no significant relationship
with trophic similarity. Considering species habitat, the
``Benthopelagic'' and ``Pelagic'' categories presented the two species
with the highest mean interaction strength (Orcinus orca and
Mesonychoteuthis hamiltoni, respectively), though ``Demersal'' and
``Land-based'' habitats concentrated the majority of the species with
relatively higher interaction strength. Species inhabiting the benthic
realm showed the lowest mean interaction strength (Figure 4D).

@@

It's important to note that since the proportion of Jacobians that are
locally stable or QSS was zero for the Weddell Sea food web, we
considered the mean maximum eigenvalue as the stability index. When
performing the extinction simulations, our results showed that the
majority of the species had no significant impact on the stability of
the food web when extinct (Figure 5). This is shown in Figure 5, where
most of the points lie around the zero value of the `Stability
difference' y axis, meaning that the stability for the food web minus a
given species is similar (not significant) to the stability for the
whole network. However, there were few species that when become extinct
altered significantly the stability of the food web. Most of such
species generated positive significant differences in the stability of
the food web, which mean that they increased the network's stability
when deleted. There were two exceptions, Pagetopsis macropterus
(demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod),
that when deleted decreased the stability. Overall, 15 out of 490
species (3.06\%) gave rise to significant changes in the food web's
stability (Table 2).

After exploring the stability difference against the species properties
(Figure 5), we found that those species that generated a significant
impact on the stability of the food web were characterized by: 1) high
mean interaction strength; 2) mid to high trophic levels (TL
\textgreater{} 3.2); 3) relatively high number of interactions (Degree
\textgreater{} 25); and 4) mid to low trophic similarity (TS \textless{}
0.16). Habitat wise, species with a significant impact on the stability
were present in all habitats, except for the benthic realm. Table 2
shows these results for such species.

@@

Principal results: 1) First study to * consider the interaction
dimensionality to estimate interaction strength in a food web framework

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\item
  Asymmetric distribution of IS at the food web level; dominance of weak
  interactions
\item
  Strong relationship between interaction strength and trophic level.
  Less stronger but evident relationship with degree
\item
  Few species are important in terms of stability: discuss species
  characteristics
\item
\item
  Dominance of weak interactions in concordance with other studies.
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
\item
  Our results showed that only few species are key in terms of stability
  in the Weddell Sea food web. This is in concordance with other studies
  of complex empirical food webs in high-latitude marine ecosystems
  \citep{Kortsch2015, Marina2018, Rodriguez2022}. Such key species are
  characterized by: 1) high mean interaction strength; 2) mid to high
  trophic level; 3) relatively high number of interactions; and 4) mid
  to low trophic similarity.
\end{enumerate}

Our results contradict studies suggesting that mid-trophic level species
are key in marine polar ecosystems.

The trophic similarity index might be interpreted as a proxy for the
functional redundancy of the food web. In this sense, \ldots{} Low
functional redundancy at key trophic levels makes these ecosystems
(polar pelagic) particularly sensitive to change. \citep{Murphy2016}

@@

\begin{figure}
\includegraphics[width=12cm]{Fig3_IntDist} \caption{Frequency distribution of interaction strengths for the Weddell Sea food web. Total number of interactions = 16041.}\label{fig:unnamed-chunk-3}
\end{figure}
